Highly Available Transactions: Virtues and Limitations

### 2013-05-01 writeup/hat-analysis.tex
PDB pass

As Brewer states, ``systems and database communities are separate but
overlapping (with distinct vocabulary)''~\cite{brewer-slides}. With
this challenge in mind, when possible, we build on existing properties
and definitions from the database and distributed systems literature,
providing a brief, informal explanation and example for each
guarantee. The database isolation guarantees require particular care,
since different DBMSs often use the same terminology for different
mechanisms and may provide additional guarantees in addition to our
implementation-agnostic definitions.  We draw largely on Atul Adya's
dissertation~\cite{adya} and somewhat on its predecessor work: the
ANSI SQL specification~\cite{ansi-sql} and Berenson et al.'s
subsequent 1995 critique~\cite{ansicritique}.

@@

To begin, \textbf{Read Uncommitted} isolation is captured by Adya as
\textit{PL-1}. In this model, writes to each object are totally
ordered, corresponding to the order in which they are installed in the
database. In a distributed database, different replicas may receive
writes to their local copies of data at different times, but the order
in which replicas overwrite versions should be consistent with the total order for
each item. \textit{PL-1} requires that writes to different objects are
ordered consistently across transactions, prohibiting Adya's
phenomenon $G0$, also called ``Dirty Writes''~\cite{adya}. If we
create a graph of transactions with edges from one transaction to
another if the former overwrites the latter's write to the same
object, then, under Read Uncommitted, there should be no cycles in
this graph. Consider the following example:
\begin{align*}
\small\vspace{-1em}
T_1 &: w_x(1)~w_y(1)
\\T_2 &: w_x(2)~w_y(2)
\end{align*}
In this example, under Read Uncommitted, it is impossible for the
database to order $T_1$'s $w_x(1)$ before $T_2$'s $w_x(2)$ but order
$T_2$'s $w_y(2)$ before $T_1$'s $w_y(1)$.  Read Uncommitted is easily
achieved via timestamping each update in a transaction with the same
logical timestamp (which is unique across transactions; e.g., client
ID and logical clock) and applying a ``last writer wins'' conflict
reconciliation policy at each replica. Later properties will
strengthen Read Uncommitted.

\textbf{Read Committed} isolation is particularly important in
practice as it is the default of many DBMSs. Centralized
implementations differ, with some based on long-duration exclusive
locks and short-duration read locks~\cite{gray-isolation} and others
based on multiple versions. The implementations often provide recency
and monotonicity properties beyond the simple meaning from the name,
which is what is expressed in the implementation-agnostic definition:
under Read Committed, transactions should not access uncommitted or
intermediate versions of data items. This prohibits both ``Dirty
Writes'', as above, and also ``Dirty Reads'' phenomena.  This
isolation is Adya's \textit{PL-2} and is formalized by prohibiting
Adya's \textit{G1\{a-c\}} (or ANSI's $P1$, or ``broad'' $P1$ [2.2]
from Berenson et al.). For instance, in the example below, $T_3$
should never see $a=1$, and, if $T_2$ aborts, $T_3$ should never read
$a=3$:
\begin{align*}
\small\vspace{-1em}
T_1 &: w_x(1)~w_x(2)
\\T_2 &: w_x(3)\\
T_3 &: r_x(a)\vspace{-1em}
\end{align*}
It is fairly easy for a HAT system to prevent ``Dirty Reads'': if each
client never writes uncommitted data to shared copies of the database, then
transactions will never read each others' dirty data. As a simple
solution, clients can buffer their writes until they commit, or,
alternatively, can send them to servers, who will not deliver their
value to other readers until notified that the writes have been
committed. This implementation does not provide recency or
monotonicity guarantees but satisfies the implementation-agnostic
definition.

Several different properties have been labeled \textbf{Repeatable
  Read} Isolation. As we will show in
Section~\ref{sec:unachievable-acid}, some of these are not achievable
in a HAT system. However, the ANSI standardized
implementation-agnostic definition~\cite{ansi-sql} \textit{is}
achievable and directly captures directly the spirit of the term: if a
transaction reads the same data twice, it sees the same value each
time (preventing ``Fuzzy Read,'' or $P2$). In this paper, to
disambiguate between other definitions of ``Repeatable Read,'' we will
call this property ``cut isolation,'' since each transaction reads
from a non-changing cut, or snapshot, over the data items. If this
property holds over reads from discrete data items, we call it
\textbf{Item Cut Isolation (I-CI)}, and if we also expect a cut over
predicate-based reads (preventing Phantoms~\cite{gray-isolation}, or
Berenson et al.'s $P3/A3$~\cite{ansicritique}), we have the stronger
property of \textbf{Predicate Cut-Isolation}. In the example below,
under both levels of cut isolation, $T_3$ must read $a=1$:
\begin{align*}
\small
T_1 &: w_x(1)
\\T_2 &: w_x(2)
\\T_3 &: r_x(1)~r_x(a)
\end{align*}
It is possible to satisfy Item Cut Isolation with high availability by
having clients cache their reads such that the value that they read
never changes unless they overwrite it themselves. The cache can be
cleared at the end of each transaction and can alternatively be
accomplished on (sticky) replicas via multi-versioning. Predicate Cut
Isolation is also achievable in HAT systems via similar caching or
multi-versioning that tracks predicate ranges instead of item-based
reads.

@@

\textbf{Transactional atomicity (TA)} is core to ACID guarantees. Although, at
least by the ACID acronym, it is not an ``isolation'' property, TA
restricts transactions' ability to view the effects of partially
completed transactions. Under TA, once some of the effects of a
transaction $T_i$ are observed by another transaction $T_j$,
thereafter, all effects of $T_i$ are observed by $T_j$. Together with
item cut isolation, TA prevents Read Skew anomalies (Berenson et al.'s
A5A~\cite{ansicritique}). As an example of TA, because $T_2$ has read
$T_1$'s write to $y$, $T_2$ must observe $b=c=1$ (or later versions
for each key):
\vspace{-1em}
\begin{align*}
\small
T_1 &: w_x(1)~w_y(1)~w_z(1)
\\T_2 &: r_x(a)~r_y(1)~r_x(b)~r_z(c)~\\[-1.5em]
\end{align*}
$T_2$ can also observe $a=\bot$, $a=1$, or a later version of
$a$. Notably, TA requires Read Committed isolation: observing all
effects of a transaction implicitly requires observing the final
(committed) effects of a transaction as well.

@@

We sketch a non-sticky TA algorithm and provide greater detail in our
technical report~\cite{hat-tr}. Each replica waits to reveal write $w$
from transaction $T$ until all of the writes in $T$ are present on all
respective replicas (are \textit{pending stable}). Once they are,
replicas asynchronously reveal their writes. To prevent clients from
reading writes that have been revealed on some servers but not others,
clients can attach a vector that uniquely identifies the transaction:
this can be the transaction ID and a list of keys in the
transaction. When clients attempt to read a data item from a replica,
they can send the highest ID they have observed for that item;
replicas respond with either a matching write (which may or may not be
pending stable) or a write with higher ID that is pending
stable. Clients can reset their record of IDs and keys upon
transaction commit. Servers keep two sets of versions for each data
item: the value with the highest ID that is pending stable and a set
of values that are not yet pending stable.

This non-sticky implementation is entirely masterless and neither reads
nor writes block for coordination. Moreover, there are several
possible optimizations that, for brevity, we do not describe
here. However, one key property is that we do not, in general, dictate
when writes become visible. Accordingly, it is highly
available. However, we can optimize write visibility delay by ensuring
that clients are sticky with disjoint groups of servers. Once all
transaction values are \texttt{pending} within a group of servers,
values can be made \texttt{stable} (as opposed to once they are
present on all servers). For example, in a multi-datacenter setting,
all clients and servers within a datacenter may form a sticky group.

@@

\vspace{.5em}\noindent{\textbf{Durability}} A client requiring that its
transactions' effects survive $F$ server faults requires at least
$F+1$ non-failing replicas; this affects availability.

>>>>

As Brewer states, ``systems and database communities are separate but
overlapping (with distinct vocabulary)''~\cite{brewer-slides}. With
this challenge in mind, we build on existing properties and
definitions from the database and distributed systems literature,
providing a brief, informal explanation and example for each
guarantee. The database isolation guarantees require particular care,
since different DBMSs often use the same terminology for different
mechanisms and may provide additional guarantees in addition to our
implementation-agnostic definitions.  We draw largely on Atul Adya's
dissertation~\cite{adya} and somewhat on its predecessor work: the
ANSI SQL specification~\cite{ansi-sql} and Berenson et al.'s
subsequent critique~\cite{ansicritique}.

@@

To begin, \textbf{Read Uncommitted} isolation is captured by Adya as
\textit{PL-1}. In this model, writes to each object are totally
ordered, corresponding to the order in which they are installed in the
database. In a distributed database, different replicas may receive
writes to their local copies of data at different times, but the order
in which replicas overwrite versions should be consistent with the total order for
each item. \textit{PL-1} requires that writes to different objects are
ordered consistently across transactions, prohibiting Adya's
phenomenon $G0$, also called ``Dirty Writes''~\cite{adya}. If we
create a graph of transactions with edges from one transaction to
another if the former overwrites the latter's write to the same
object, then, under Read Uncommitted, there should be no cycles in
this graph. Consider the following example:
\begin{align*}
\small\vspace{-1em}
T_1 &: w_x(1)~w_y(1)
\\T_2 &: w_x(2)~w_y(2)
\end{align*}
In this example, under Read Uncommitted, it is impossible for the
database to order $T_1$'s $w_x(1)$ before $T_2$'s $w_x(2)$ but order
$T_2$'s $w_y(2)$ before $T_1$'s $w_y(1)$. Read Uncommitted is easily
achieved via timestamping each update in a transaction with the same
logical timestamp (which is unique across transactions; e.g.,
combining a client ID and logical clock) and applying a ``last writer
wins'' conflict reconciliation policy at each replica. Later
properties will strengthen Read Uncommitted.

\textbf{Read Committed} isolation is particularly important in
practice as it is the default of many DBMSs. Centralized
implementations differ, with some based on long-duration exclusive
locks and short-duration read locks~\cite{gray-isolation} and others
based on multiple versions. The implementations often provide recency
and monotonicity properties beyond the simple meaning from the name,
which is what is expressed in the implementation-agnostic definition:
under Read Committed, transactions should not access uncommitted or
intermediate versions of data items. This prohibits both ``Dirty
Writes'', as above, and also ``Dirty Reads'' phenomena.  This
isolation is Adya's \textit{PL-2} and is formalized by prohibiting
Adya's \textit{G1\{a-c\}} (or ANSI's $P1$, or ``broad'' $P1$ [2.2]
from Berenson et al.). For instance, in the example below, $T_3$
should never see $a=1$, and, if $T_2$ aborts, $T_3$ should never read
$a=3$:
\begin{align*}
\small\vspace{-1em}
T_1 &: w_x(1)~w_x(2)
\\T_2 &: w_x(3)\\
T_3 &: r_x(a)\vspace{-1em}
\end{align*}
It is fairly easy for a HAT system to prevent ``Dirty Reads'': if each
client never writes uncommitted data to shared copies of data, then
transactions will never read each others' dirty data. As a simple
solution, clients can buffer their writes until they commit, or,
alternatively, can send them to servers, who will not deliver their
value to other readers until notified that the writes have been
committed. This implementation does not provide recency or
monotonicity guarantees but satisfies the implementation-agnostic
definition.

Several different properties have been labeled \textbf{Repeatable
  Read} Isolation. As we will show in
Section~\ref{sec:unachievable-acid}, some of these are not achievable
in a HAT system. However, the ANSI standardized
implementation-agnostic definition~\cite{ansi-sql} \textit{is}
achievable and directly captures directly the spirit of the term: if a
transaction reads the same data twice, it sees the same value each
time (preventing ``Fuzzy Read,'' or $P2$). In this paper, to
disambiguate between other definitions of ``Repeatable Read,'' we will
call this property ``cut isolation,'' since each transaction reads
from a non-changing cut, or snapshot, over the data items. If this
property holds over reads from discrete data items, we call it
\textbf{Item Cut Isolation (I-CI)}, and if we also expect a cut over
predicate-based reads (e.g., \texttt{SELECT WHERE}; preventing
Phantoms~\cite{gray-isolation}, or Berenson et al.'s
$P3/A3$~\cite{ansicritique}), we have the stronger property of
\textbf{Predicate Cut-Isolation}. In the example below, under both
levels of cut isolation, $T_3$ must read $a=1$:
\begin{align*}
\small
T_1 &: w_x(1)
\\T_2 &: w_x(2)
\\T_3 &: r_x(1)~r_x(a)
\end{align*}
It is possible to satisfy Item Cut Isolation with high availability by
having clients cache their reads such that the value that they read
never changes unless they overwrite it themselves. The cache can be
cleared at the end of each transaction and can alternatively be
accomplished on (sticky) replicas via multi-versioning. Predicate Cut
Isolation is also achievable in HAT systems via similar caching or
multi-versioning that tracks entire logical ranges of predicates
in addition to item-based reads.

@@

\textbf{Transactional atomicity (TA)} is core to ACID guarantees. Although, at
least by the ACID acronym, it is not an ``isolation'' property, TA
restricts transactions' ability to view the effects of partially
completed transactions. Under TA, once some of the effects of a
transaction $T_i$ are observed by another transaction $T_j$,
thereafter, all effects of $T_i$ are observed by $T_j$. Together with
item cut isolation, TA prevents Read Skew anomalies (Berenson et al.'s
A5A~\cite{ansicritique}). As an example of TA, because $T_2$ has read
$T_1$'s write to $y$, $T_2$ must observe $b=c=1$ (or later versions
for each key):
\begin{align*}
\small
T_1 &: w_x(1)~w_y(1)~w_z(1)
\\T_2 &: r_x(a)~r_y(1)~r_x(b)~r_z(c)~\\[-1.5em]
\end{align*}
$T_2$ can also observe $a=\bot$, $a=1$, or a later version of
$a$. Notably, TA requires Read Committed isolation: observing all
effects of a transaction implicitly requires observing the final
(committed) effects of a transaction as well.

@@

We sketch a non-sticky TA algorithm and provide greater detail in our
technical report~\cite{hat-tr}. Each replica waits to reveal write $w$
from transaction $T$ until all of the writes from $T$ are present on
all respective replicas (are \textit{pending stable}). Once they are,
replicas asynchronously reveal their respective writes to readers. To
prevent clients from reading incomplete portions of a transaction's
writes as they are being revealed (e.g., a client reads a data item
that from a replica that has detected pending stability, then attempts
to read another item from another replica which has not yet detected
pending stability), clients include additional metadata with each
write: a list of items in the transaction along with the transaction's
timestamp (from Read Uncommitted). When a client reads a write, the
write's list of items and timestamp form a lower bound on the versions
that the client should read for other items. When a client performs
subsequent reads, it includes a timestamp with its request that
corresponds current lower bound on reads from the respective item
Replicas respond with either a write matching the timestamp or a write
with a higher timestamp that is pending stable. Clients discard any
metadata upon transaction commit, and servers keep two sets of writes
for each data item: the write with the highest timestamp that is
pending stable and a set of writes that are not yet pending stable.

This non-sticky implementation is entirely masterless and neither
reads nor writes block for coordination. Moreover, there are several
possible optimizations that, for brevity, we do not describe
here. However, one key property is that we do not, in general, dictate
when writes become visible. Accordingly, the algorithm is highly
available. We can optimize write visibility by sticking clients are
with disjoint groups of servers: once all of a transaction's writes
are pending stable within a group of servers, writes can be made
visible (as opposed to once they are pending stable with respect to
all servers). For example, in a multi-datacenter setting, all clients
and servers within a datacenter may form a sticky group.

@@

%\vspace{.5em}\noindent{\textbf{Durability}} A client requiring that its
%transactions' effects survive $F$ server faults requires at least
%$F+1$ non-failing replicas; this affects availability.

### 2013-05-01 writeup/hat-vldb.tex
PDB pass

\begin{abstract}
To minimize network latency and remain online during server failures
and network partitions, many modern distributed data storage systems
eschew transactional functionality, which provides strong semantic
guarantees for groups of multiple operations over multiple data
items. In this work, we consider the problem of providing Highly
Available Transactions (HATs): transactional guarantees that do not
suffer unavailability during system partitions or incur high network
latency.  We introduce a new taxonomy of highly available systems and
analyze existing ACID and distributed consistency guarantees to
identify which can and cannot be achieved in HAT systems. This unifies
the literature on weak isolation, data consistency, and highly
available systems. We analytically and experimentally quantify the
availability and performance benefits of HATs---often several orders
of magnitude over wide-area networks---as well as their necessary
semantic compromises.
\end{abstract}\vspace{-.5em}

>>>>

\begin{abstract}
\noindent To minimize network latency and remain online during server
failures and network partitions, many modern distributed data storage
systems eschew transactional functionality, which provides strong
semantic guarantees for groups of multiple operations over multiple
data items. In this work, we consider the problem of providing Highly
Available Transactions (HATs): transactional guarantees that do not
suffer unavailability during system partitions or incur high network
latency.  We introduce a new taxonomy of highly available systems and
analyze existing ACID and distributed consistency guarantees to
identify which can and cannot be achieved in HAT systems. This unifies
the literature on weak isolation, data consistency, and highly
available systems. We analytically and experimentally quantify the
availability and performance benefits of HATs---two to three orders of
magnitude over wide-area networks---as well as their necessary
semantic compromises.
\end{abstract}\vspace{-.5em}

### 2013-05-01 writeup/intro.tex
PDB pass

Our main contributions in this paper are as follows. We fill the gap
between the plethora of previously proposed isolation and consistency
models and the goal of high availability.  We classify which among the
wide array of models are achievable with high availability, denoting
them as {\em Highly Available Transactions} (HATs). In doing so, we
demonstrate that although many implementations of HAT semantics are
not highly available, this is an artifact of the implementations
rather than an inherent property of the semantics. Our investigation
shows that, besides serializability, Snapshot Isolation and Repeatable
Reads are not HAT-compliant, while most other isolation levels are
achievable with high availability. We also demonstrate that many weak
data consistency models from distributed systems are also both
HAT-compliant and simultaneously achievable with traditional ACID
properties.

Our investigation is based on both impossibility results and several
constructive proof-of-concept algorithms. For example, Snapshot
Isolation (SI) and Repeatable Reads (RR) are not HAT-compliant because
they require detecting conflicts between concurrent updates (as needed
for preventing Lost Updates or Write Skew phenomena), which we prove
is unavailable. However, Read Committed isolation, Transactional
Atomicity, and many of the weaker data consistency models from
database and distributed systems are achievable via novel algorithms
that rely on multi-versioning and limited client-side caching. For
several guarantees, such as causal consistency with phantom prevention
and ANSI Repeatable Read, we consider a modified form of high
availability in which clients ``stick to'' (i.e., have affinity with)
at least one server (a property which is often implicit in the
distributed systems literature~\cite{herlihy-art, eiger, cac} but
which requires explicit consideration in a client-server replicated
database context). This sticky availability is widely
employed~\cite{eiger, vogels-defs} but is less restrictive (and
therefore more easily achievable) than traditional high availability.

>>>>

Our main contributions in this paper are as follows. We fill the gap
between the plethora of previously proposed isolation and consistency
models and the goal of high availability.  We classify which among the
wide array of models are achievable with high availability, denoting
them as {\em Highly Available Transactions} (HATs). In doing so, we
demonstrate that although many implementations of HAT semantics are
not highly available, this is an artifact of the implementations
rather than an inherent property of the semantics. Our investigation
shows that, besides serializability, Snapshot Isolation and Repeatable
Read isolation are not HAT-compliant, while most other isolation
levels are achievable with high availability. We also demonstrate that
many weak data consistency models from distributed systems are also
both HAT-compliant and simultaneously achievable with traditional ACID
properties.

Our investigation is based on both impossibility results and several
constructive, proof-of-concept algorithms. For example, Snapshot
Isolation and Repeatable Read isolation are not HAT-compliant because they
require detecting conflicts between concurrent updates (as needed for
preventing Lost Updates or Write Skew phenomena), which we prove is
unavailable. However, Read Committed isolation, Transactional
Atomicity, and many of the weaker data consistency models from
database and distributed systems are achievable via novel algorithms
that rely on multi-versioning and limited client-side caching. For
several guarantees, such as causal consistency with phantom prevention
and ANSI Repeatable Read, we consider a modified form of high
availability in which clients ``stick to'' (i.e., have affinity with)
at least one server (a property which is often implicit in the
distributed systems literature~\cite{herlihy-art, eiger, cac} but
which requires explicit consideration in a client-server replicated
database context). This sticky availability is widely
employed~\cite{eiger, vogels-defs} but is less restrictive (and
therefore more easily achievable) than traditional high availability.

### 2013-05-01 writeup/model.tex
PDB pass

Traditionally, a system provides {\em high availability} if every user
that can contact a correct (non-failing) server eventually receives a
response from that server, even in the presence of arbitrary,
indefinitely long network partitions between
servers~\cite{gilbert-cap}.\footnote{Under this definition, systems
  that require a majority of servers to be online or that require the
  database to function correctly with high (but not perfect)
  probability are not available.} As in a standard distributed
database, designated servers might perform operations for different
data items; a server that can handle an operation for a given data
item is called a \textit{replica} for that item.\footnote{There is a
  further distinction between a \textit{fully replicated} system, in
  which all servers are replicas for all data items, and a
  \textit{partially replicated} system, in which at least one server
  acts as a replica for only a subset of all data items. For
  generality, and, for applicability to many modern ``sharded'' or
  ``partitioned'' data storage systems~\cite{dynamo, pnuts, bigtable,
    spanner, hstore}, we consider partially replicated systems.} While
this standard model is useful, there are several other cases we should
consider.

Many distributed algorithms assume a model in which clients always
contact the same (sets of) server(s) across subsequent operations. As
we will discuss in Section~\ref{sec:hats}, clients can ensure
continuity between operations (e.g., reading their prior updates to a
data item) by maintaining affinity or ``stickiness'' with a server or
set of servers~\cite{vogels-defs}. Simultaneously maintaining
availability and stickiness requires fate-sharing between the client
and its server: if a client's sticky servers fail or if the client is
partitioned from its sticky servers, then it may face
unavailability. We say that a system provides \textit{sticky
  availability} if, whenever a client's operation is executed against
a correct server that has observed all of its prior operations, it
eventually receives a response, even in the presence of indefinitely
long partitions. A client may choose to become sticky available by
acting as a server itself; for example, a client might cache its reads
and writes~\cite{bolton, sessionguarantees, swift}. Any guarantee
achievable in a highly available system is achievable in a sticky high
availability system but not vice-versa.

@@

If a client wishes to execute a transaction that performs operations
on multiple data items, then the client must be able to contact and
receive a response from at least one replica for each data item. This
may result in ``lower availability'' than a non-transactional
requirement. Additionally, given the possibility of system-initiated
aborts, we need to ensure useful forward progress: a system can
trivially guarantee clients a response by always aborting all
transactions. However, this is an unsatisfactory system because
nothing good (transaction commit) ever happens; we should require a
\textit{liveness} property~\cite{transaction-liveness}. We cannot
guarantee that every transaction will commit---transactions may choose
to abort themselves---but we need to make sure that the system will
not indefinitely abort transactions on its own volition. We call a
transaction abort due to a transaction's own choosing (e.g., as an
operation of the transaction itself or due to a would-be declared
integrity constraint violation) an \textit{internal abort} and an
abort due to system implementation an \textit{external abort}.

>>>>

Traditionally, a system provides {\em high availability} if every user
that can contact a correct (non-failing) server eventually receives a
response from that server, even in the presence of arbitrary,
indefinitely long network partitions between
servers~\cite{gilbert-cap}.\footnote{Under this definition, systems
  that require a majority of servers to be online or that require the
  database to function correctly with high (but not perfect)
  probability are not available.} As in a standard distributed
database, designated servers might perform operations for different
data items; a server that can handle an operation for a given data
item is called a \textit{replica} for that item.\footnote{There is a
  further distinction between a \textit{fully replicated} system, in
  which all servers are replicas for all data items, and a
  \textit{partially replicated} system, in which at least one server
  acts as a replica for only a subset of all data items. For
  generality, and, for applicability to many modern ``sharded'' or
  ``partitioned'' data storage systems~\cite{dynamo, pnuts, bigtable,
    spanner, hstore}, we consider partially replicated systems.}

In addition to high availability, which allows operations on any replica,
distributed algorithms often assume a model in which clients always
contact the same (sets of) replica(s) across subsequent operations. As
we will discuss in Section~\ref{sec:hats}, clients can ensure
continuity between operations (e.g., reading their prior updates to a
data item) by maintaining affinity or ``stickiness'' with a server or
set of servers~\cite{vogels-defs}. Simultaneously maintaining
availability and stickiness requires fate-sharing between the client
and its server: if a client's sticky servers fail or if the client is
partitioned from its sticky servers, then it may face
unavailability. We say that a system provides \textit{sticky
  availability} if, whenever a client's operation is executed against
a correct server that has observed all of its prior operations, it
eventually receives a response, even in the presence of indefinitely
long partitions. A client may choose to become sticky available by
acting as a server itself; for example, a client might cache its reads
and writes~\cite{bolton, sessionguarantees, swift}. Any guarantee
achievable in a highly available system is achievable in a sticky high
availability system but not vice-versa.

@@

If a client wishes to execute a transaction that performs operations
on multiple data items, then the client must be able to contact and
receive a response from at least one replica for each data item. This
may result in ``lower availability'' than a non-transactional
requirement. Additionally, given the possibility of system-initiated
aborts, we need to ensure useful forward progress: a system can
trivially guarantee clients a response by always aborting all
transactions. However, this is an unsatisfactory system because
nothing good (transaction commit) ever happens; we should require a
\textit{liveness} property~\cite{transaction-liveness}. We cannot
guarantee that every transaction will commit---transactions may choose
to abort themselves---but we need to make sure that the system will
not indefinitely abort transactions on its own volition. We call a
transaction abort due to a transaction's own choosing (e.g., as an
operation of the transaction itself or due to a would-be declared
integrity constraint violation) an \textit{internal abort} and an
abort due to system implementation or operation an \textit{external abort}.

### 2013-05-01 writeup/motivation.tex
PDB pass

These results---which do not even consider server failures---indicate
that network partitions \textit{do} occur within and across modern
datacenters. We observe that, as a general trend, partitions exist and
must accordingly be met with either unavailability at some servers or,
as we will discuss, relaxed semantic guarantees.

@@

Even with fault-free networks, distributed systems face the challenge
of communication latency, Deutsch's second ``Fallacy.'' In this
section, we quantify round-trip latencies, which are often
large---hundreds of milliseconds---in a geo-replicated,
multi-datacenter context.  Fundamentally, the speed at which two
servers can communicate is (according to modern physics) bounded by
the speed of light. In the best case, two servers on opposite sides of
the Earth---communicating via a hypothetical link traversing the
planet's diameter---require a minimum 85.1ms RTT (133.7ms if sent at
surface level). As services are replicated to multiple, geographically
distinct sites, this cost of communication increases.

@@

In real deployments, messages travel slower than the speed of light
due to routing, congestion, and server-side overheads. To illustrate
the difference between intra-datacenter, inter-datacenter, and
inter-planetary networks, we performed a measurement study of network
behavior on Amazon's EC2, a widely-used public compute cloud. We
measured one week of ping times between all seven EC2 geographic
``regions'', across three ``availability zones'' (closely co-located
datacenters), and within a single ``availabilty zone'' (datacenter),
at a granularity of 1s (dataset to be released). We summarize the
results of our network measurement study in Table~\ref{table:rtt}. On
average, intra-datacenter communication (Table 1a) is between 1.82 and
6.38 times faster than across geographically co-located datacenters
(Table 1b) and between 40 and 647 times faster than across
geographically distributed datacenters (Table 1c). The cost of
wide-area communication exceeds the speed of light: for example, while
a speed-of-light RTT from S\~{a}o Paulo to Singapore RTT is 106.7ms,
ICMP packets incur an average 362.8ms RTT (95th percentile: 649ms). As
shown in Figure~\ref{fig:rtt}, the distribution of latencies varies
between links, but the trend is clear: remote communication has a
substantial cost.

\section{When is ``ACID'' ACID?}
\label{sec:modernacid}

>>>>

These results---which do not consider server failures---indicate
that network partitions \textit{do} occur within and across modern
datacenters. We observe that, as a general trend, partitions exist and
must accordingly be met with either unavailability at some servers or,
as we will discuss, relaxed semantic guarantees.

@@

Even with fault-free networks, distributed systems face the challenge
of communication latency, Deutsch's second ``Fallacy.'' In this
section, we quantify round-trip latencies, which are often
large---hundreds of milliseconds---in a geo-replicated,
multi-datacenter context.  Fundamentally, the speed at which two
servers can communicate is (according to modern physics) bounded by
the speed of light. In the best case, two servers on opposite sides of
the Earth---communicating via a hypothetical link along the planet's
diameter---require a minimum 85.1ms RTT (133.7ms if sent at
surface level). As services are replicated to multiple, geographically
distinct sites, this cost of communication increases.

@@

In real deployments, messages travel slower than the speed of light
due to routing, congestion, and server-side overheads. To illustrate
the difference between intra-datacenter, inter-datacenter, and
inter-planetary networks, we performed a measurement study of network
behavior on Amazon's EC2, a widely-used public compute cloud. We
measured one week of ping times between all seven EC2 geographic
``regions'', across three ``availability zones'' (closely co-located
datacenters), and within a single ``availabilty zone'' (datacenter),
at a granularity of 1s (dataset to be released). We summarize the
results of our network measurement study in Table~\ref{table:rtt}. On
average, intra-datacenter communication (Table 1a) is between 1.82 and
6.38 times faster than across geographically co-located datacenters
(Table 1b) and between 40 and 647 times faster than across
geographically distributed datacenters (Table 1c). The cost of
wide-area communication exceeds the speed of light: for example, while
a speed-of-light RTT from S\~{a}o Paulo to Singapore RTT is 106.7ms,
ping packets incur an average 362.8ms RTT (95th percentile: 649ms). As
shown in Figure~\ref{fig:rtt}, the distribution of latencies varies
between links, but the trend is clear: remote communication has a
substantial cost.

\section{ACID in the Wild}
\label{sec:modernacid}
